---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan. 27

The command to set a working directory is setwd()

Also, this is just text. If you want to run code, you can like this:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```
## <dbl> means "number" for R
-----
## Sorting Data (Descending - Biggest to Smallest)

To sort a certain dataset by descending order, use this (replace assets with appropriate ones, but use the commands)

This first groups the data by county name, then counts the number of loans, then arranges them in descending order

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
----------------------------------
## Saving something as a variable
Take a look at the following code, which allows you to save a set of code as a variable (which then allows it to be viewed in the Data Viewer)

```{r}
wv_summary <- wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```
---------------------------------
# A good way to introduce yourself to the contents of you data set:

```{r}
head(maryland_ppp)
summary(maryland_ppp)
colnames(maryland_ppp)
glimpse(maryland_ppp)
```

"head":
"summary:"
"colnames:"
"glimpse:"

-------------------------
### Mutate
**Task** Run the following code to create a new column called `percent_payroll` based on a calculation using two existing columns. This allows you to add a new column (which in this example, we name "percent_payroll"), or change an existing column.

```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = payroll_proceed/amount
)
```
-------------------
### SELECT: Better percentage calculation
**Task** Run the following code to make our new column called `percent_payroll` show a percentage instead of a decimal. Doesn't make the rest of your data disappear, but it shows you specific data you want to see.

```{r}
# select a range of data, then make it a percentage
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)
```
-------------------
### ARRANGE: Mutate with ordering
**Task** Run the following code to order by our new column. By adding the "%>% arrange(desc(percent_payroll))", we can arrange the percent_payroll data in DESCENDING ORDER

```{r}
# better ordering?
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(desc(percent_payroll))
```
----------------
### To organize data CONDITIONALLY, using "If/Else" and "Filter"
One question you might have is whether the lenders servicing these loan applications are located in Maryland or outside. We can add a column using mutate and populate it based on the state of the lender.

**Task** Run the following code to create a new column called `in_out` based on the values of `servicing_lender_state`. Add a description of what this code does to your reference notebook.

```{r}
# if/else example. This creates a "yes or no" answer, by creating new columns. This is good for a binary question, when an element of data is either one thing or another, and not a third (or fourth, fifth, etc.) thing
maryland_with_in_out <- maryland_ppp %>%
  mutate(
    #create a new column "in_out", which uses the "if_else()" function to populate "in_out" if the value of "servicing_lender_state" is in Maryland. If it is, then the column will say "IN." If it's not in Maryland, then the new column will say "OUT"
    in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
      )
  )
```

For a single variable, you can use FILTER:
```{r}
maryland_ppp %>%
  # filter() filters data by a single column identifier
  filter(servicing_lender_state == 'MD')
```
###Here is a more in-depth description of using filter:

## Filters

Filter is a tidyverse verb that uses criteria you supply to return a subset of the original data. Here's how it is structured:

`filter(COLUMN_NAME OPERATOR CRITERIA)`

Where COLUMN_NAME is any column name from your dataframe, OPERATOR is some comparison operator like == or > or < and CRITERIA is something to compare the values in COLUMN_NAME to.


###This is from Lab 02, where we had to find which companies had the max jobs retained of 500, and then arrange by which loan recipient had their loan approved the earliest:

```{r}
west_virginia_ppp %>% filter(jobs_retained == 500) %>% arrange(date_approved) 
```

###This is also from Lab 02, and in answer to the question: Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?

###The best way to do this is to first filter out the unecessary data by using filter, which eliminates any undisbursed loans less than 0. The code the groups the data by the servicing lender, the servicing lender state. Then it counts the loans and arranges them from largest to smallest.

```{r}
west_virginia_ppp %>%
  filter(undisbursed_amount > 0) %>%
  group_by(servicing_lender_name, servicing_lender_state) %>%
  summarise(
    total_loans = n()
  ) %>%
  arrange(desc(total_loans))
```
-----------------------
### Task 9: Create a new column using case_when, for more multi-layered data grouping needs
One common use of mutate is to create a new column and then populate it based on values in an existing column. For example, if we wanted to put the values in `jobs_retained` into broader categories in a new column called `jobs_retained_category` and we have multiple categories. We can use a function called `case_when` to make that easier.

**Task** Run the following code to create a new column called `jobs_retained_category` based on the values of `jobs_retained`. Add a description of what this code does to your reference notebook.

```{r}
# case_when example. This creates separate "buckets" to help sort the data, decided by "case_when()." This allows you to put in a number of different scenarios: "when the number of jobs retained is less than 10, then put it in the "10_to_25" column." 

#this also saves the following data as a new variable, called "maryland_jobs_categories"
```{r}
maryland_jobs_categories <- maryland_ppp %>%
mutate(
  jobs_retained_category = case_when(
    jobs_retained < 10 ~ 'under_10',
    jobs_retained >= 10 & jobs_retained < 25 ~ '10_to_25',
    jobs_retained >= 25 & jobs_retained < 50 ~ '25_to_50',
    jobs_retained >= 50 & jobs_retained < 100 ~ '50_to_100',
    jobs_retained >= 100 & jobs_retained < 250 ~ '100_to_250',
    jobs_retained >= 250 & jobs_retained < 500 ~ '250_to_500',
    jobs_retained == 500 ~ '500'
  )
)
```
There's a lot going on here, so let's unpack it. It starts out as a typical mutate statement, but like our if_else example, case_when introduces some new things. Each line is basically a filter followed by `~` and then a value for the new column for records that match that filter. Here is [more detail](https://www.sharpsightlabs.com/blog/case-when-r/) on using `case_when`.
  
# FYI: The below code allows you to summarize the data above by grouping the data by job-retained ranges, and show how many entities retained jobs in those ranges.
maryland_jobs_categories %>% group_by(jobs_retained_category) %>% summarise(count = n())

###USE FILTER TO FIND <NA> VALUES:

```{r}
ppp_maryland_loans %>%
  group_by(forgiveness_date) %>%
  summarise(
    count=n()
  ) %>%
  filter(is.na(forgiveness_date)) %>%
  arrange(desc(count))
```

###USE FILTER TO FIND VALUES THAT ARE *NOT <NA>:
```{r}
ppp_maryland_loans %>%
  group_by(forgiveness_date) %>%
  summarise(
    count=n()
  ) %>%
  filter(is.na(forgiveness_date)) %>%
  arrange(desc(count))
```

----------

### Task 16: Use mutate and lubridate to add a column called `Month` and see which month had the most applications.
**Task** Run the following code to create a new column based on `date_approved`. 

This code creates a new column, grouped by month, and sorts the date from highest to smallest (in terms of the number of loans)

```{r}
maryland_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

###LOOK FOR DATE GAPS IN DATA:
##This code creates a new column (month_year_approved), which is where the floor_date of the (date_approved_"month") column will go. The code extracts the monthly loan numbers, groups them by the month and year approved, counts them, then arranges them in ascending order by month_year_approved).

```{r}
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)
```