---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan. 27

The command to set a working directory is setwd()

Also, this is just text. If you want to run code, you can like this:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```
## <dbl> means "number" for R
-----
## Sorting Data (Descending - Biggest to Smallest)

To sort a certain dataset by descending order, use this (replace assets with appropriate ones, but use the commands)

This first groups the data by county name, then counts the number of loans, then arranges them in descending order

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```

