---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here

install.packages('janitor')
```



## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)

wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia, got the largest _number_ of loans? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

1) Charleston, 2802

2) South Charleston, 478

3) Saint Albans, 296

4) Dunbar, 162

5) Nitro, 140


Note: To answer this question, you'll need to see every variation of the "city" column in Kanawha County, and then fully clean the city field for those cities. 

Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1:
```{r}
#Glimpse data to see what it looks like first
glimpse(wv_ppp_loans)

#The <city> column is <chr>, which is correct. But from this, I can see variations in spelling and titlecase.
```

```{r}
#the first step is to isolate the cities so I only see the ones from Kanawha County. I want to see this in a new table so I can track it.
Kanawha_Cities <- wv_ppp_loans %>%
filter(project_county_name == "KANAWHA") %>%


#now I need to clean the city names first
#Standardizes a column to one case. In this case, it's changing the "city" column to "title" case.

mutate(city = str_to_title(city)) %>%

#I can see that there are variations in some cities, like "S Charleston." The actual city name is "South Charleston." I will now fix the spellings to make them standard.
  mutate(city = case_when(
    #If the city is "Anold," change it to "Arnold"
    city == "S Charleston" ~ "South Charleston",
    city == "So Charleston" ~ "South Charleston",
    city == "Charelston" ~ "Charleston",
    city == "Balle" ~ "Belle",
    city == "Elkveiw" ~ "Elkview",
    city == "Crosslanes" ~ "Cross Lanes",
    str_detect(city,"^St") ~ "Saint Albans",
    TRUE ~ city
  )) %>%
   mutate(zip = str_sub(zip, start=1L, end=5L)) 
  

```

#I now need to identify how many unique cities there are, and how many loans each city received.
```{r}  
Clean_Cities <- Kanawha_Cities %>%
 
  group_by(city) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))
```

```{r}
Clean_Cities %>%
  distinct()
```


#now I need to group the date by city
group_by(city) %>%
  
Q2: Find the 5-digit ZIP code in West Virginia that had the highest total amount of loans (by dollar value). What is the name of the business that got the largest loan (by dollar value) in that zip code? Write up a short paragraph that also answers these questions:


* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2:

The WV Zipcode with the highest total amount of loans is 25301, which is in Charleston (with a whopping sum of ($160,721,320.88). Within that Zipcode, the business with the largest loan is "THE EYE AND EAR CLINIC OF CHARLESTON DBA CHARLESTON SURGICA," which received $1,520,007.00. Of that loan, $1,538,254.68 (or 101%) was forgiven. Since there was interest on these loans which accrued, that explains why the forgiveness amount was more than the loan amount.

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
wv_zips <- wv_ppp_loans %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  group_by(zip) %>% 
  summarise(total = sum(amount)) %>% 
arrange(desc(total))

wv_zips
```
  
  
  

# display the cleaned dataset
```{r}
wv_zips %>%
  distinct()
```
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
```{r}
wv_ppp_loans %>%
  filter(zip == 25301) %>%
  group_by(name,amount, forgiveness_amount) %>%
  arrange(desc(amount))
```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  
  3105300
  * What percentage of the total loan amount have been forgiven?
  101.3%
  
  * How did the U.S. House member vote on the legislation initially author zing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

Some hints to help you answer this question:

* West Virginia's three U.S. House members are listed [here](https://www.govtrack.us/congress/members/WV#representatives)
* To find businesses they or their families own, you can search for financial disclosure forms for U.S. House members [here](https://disclosures-clerk.house.gov/PublicDisclosure/FinancialDisclosure#Search).
  * To speed things up, here are direct links to the three U.S. House members most recent financial disclosures, which I found by searching the site linked above.
  * [Rep. 1](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042042.pdf) (GGM Consulting LLC)
  * [Rep. 2](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10040330.pdf) 
  * [Rep. 3](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042734.pdf) (DM Motors, Dutch Miller) Chevrolet, )
* When you examine the disclosure forms, there may be lots of stock sales and retirement accounts mixed in.  Ignore those. In the table in the disclosure forms, look for "Partnership" or "C-Corp" in the column named "Income Type"
* This is a link to details on passage of the [CARES ACT](https://www.congress.gov/bill/116th-congress/house-bill/748). You can find a link to how the House voted by clicking on a link from [this page](https://www.congress.gov/bill/116th-congress/house-bill/748/actions).

A3:
As it turns out, Rep. Carol Devine Miller's (R—W.V.) husband, Matt — along with two of his sons — owns a chain of car dealerships in Huntington, Barboursville, South Charleston and Charleston. Of the seven dealerships they list on their website, five received PPP loans: DM MOTOR, INC, DUTCH MILLER CHEVROLET, INC., DUTCH MILLER OF CHARLESTON, INC, DUTCH MILLER OF CHARLOTTE, INC, DUTCH MILLER SUBARU, INC. 

The sum total of the loans those five businesses received stood at $3,105,300. Not a small sum, but since 101% of the loans were forgiven (a total of $3,140,471), not a bad deal at all.

Rep. Carol Miller, who sits on the Worker and Family Support Subcommittee was a "Yea" vote on the original CARES act. This turned out to be a prescient decision, as it may well have contributed to her husband's company retaining 265 jobs, which is 88% of the 300 people they claim to employ on the company's website.


```{r}

# Hint
# First, create a dataframe that only has loans to those five companies
dutch_miller_combined_ppp <- wv_ppp_loans %>% 
  filter(str_detect(name, "DUTCH MILLER |DM MOTOR"))
```

```{r}
# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies
dutch_miller_combined_ppp %>% 
  summarise(total = sum(amount), 
          forgiven_sum = sum(forgiveness_amount),
         ((forgiven_sum/total)*100),
            (total_jobs = sum(jobs_retained)))
```

