---
title: "lab_07"
author: "sean mussenden"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)


# Remove scientific notation
options(scipen=999)

library(rvest)
library(tidyverse)
library(janitor)
```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest to scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.

A1.
*101!*

##Steps:
**Store the URL**
```{r}
criminal_fraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"
```

**Read the HTML**
```{r}
criminal_fraud_HTML <- criminal_fraud_url %>% 
  read_html
```

**Grab the elements I need**
```{r}
criminal_fraud_elements_by_name <- criminal_fraud_HTML %>% 
  html_elements('ul li b') %>% 
  html_text() %>% 
  as_tibble()
criminal_fraud_elements_by_name
```

--------
Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
A2.
*31*

```{r}
criminal_fraud_elements_by_judicial_district <- criminal_fraud_HTML %>% 
  html_elements('p b i') %>% 
  html_text() %>% 
  as_tibble()
criminal_fraud_elements_by_judicial_district 
```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
A4.
*Smith is accused of scheming to "obtain by fraud" millions of dollars in PPP loans. Smith is also accused of conspiring to expand the scheme by "recruiting other confederate loan applicants in order to receive kickbacks from those confederates." Smith was charged with wire fraud, bank fraud, and conspiracy to commit wire fraud and bank fraud, according to a DOJ press release.*

*As an editor, I would have assigned a reporter to this story when it was filed in court because of the high dollar amount, and the fact that it was part of a greater conspiracy which ended up crossing state lines. That is in the public's interest to know, and important to report.*

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? 

In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
A5.
**The DOJ has filed cases against people in 33% of judicial districts.**

##Steps:
- Load in the Wikipedia URL
- Pull in the table from the Wikipedia page, save it as an element
```{r}
##Load URL
active_circuit_courts <- "https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts" 
  active_circuit_courts_HTML <- active_circuit_courts %>% 
  read_html
```

##Pull in the table, save it as an element
```{r}
active_circuit_courts_table <- active_circuit_courts_HTML %>% 
  html_table()
active_circuit_courts_table
```
##Find the percentage
```{r}
(31/94)*100
```



Q6. What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
A6.
**In the press release, the DOJ representative explains the complexity of these investigations and the fact that each investigation includes the cooperation of local partners and law enforcement. So that makes sense as for the pace of the investigations (which the DOJ celebrates as remarkable, given the amount of data).**

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://dwillis.github.io/jour472files/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://dwillis.github.io/jour472files/ppp_scraping_example/states/alabama.html). 

You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question. 

##Don't forget to calculation the loans per 100,000.##

A7.

##Combine the URLs into the table
```{r}
# Define url of page we want to scrape

ppp_by_state_url <- "https://dwillis.github.io/jour472files/ppp_scraping_example/index.html"

# Read in all html from table, store all tables on page as nested list of dataframes.
states_and_urls <- ppp_by_state_url%>%
  read_html() %>%
  html_table()

# show the dataframe
states_and_urls
```


