---
title: "lab_04"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
# add code here to accomplish the task
maryland_ppp <- read_rds("lab_04.rds")
maryland_ppp
```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the pre-lab. There's a particular set of steps you'll need to perform to make that work properly, so take your time and check your work.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
# add code here to accomplish the task
naics_codes <- read_csv("data/naics_codes.csv")
naics_codes
```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.


```{r}
# add code here to accomplish the task
maryland_ppp_with_naics <- maryland_ppp %>% left_join(naics_codes) %>% 
  mutate(zip5 = str_sub(zip, 1, 5))
```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one.

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character. IMPORTANT: you can see what the column names represent by checking the data documentation, which can be found here: https://geodata.md.gov/imap/rest/services/Demographics/MD_CensusData/FeatureServer/1. You will need to refer to this to answer the questions below.

```{r}
# load in the ZCTA data
md_ZCTA5N <- read_csv('data/maryland_zcta.csv')
md_ZCTA5N
```

```{r}
#change the ZCTA5N from <dbl> to <chr>
maryland_zcta_clean <- md_ZCTA5N %>% mutate(across(ZCTA5N, as.character))
maryland_zcta_clean
```
Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
# add code here to accomplish the task
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta_clean, by=c("zip5"="ZCTA5N"))
```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?

**A1** [write your answer here]
1, Taxi Service
2, Beauty Salons
3, Full-Service Restaurants
4, Offices of Real Estate Agents and Brokers
5, All Other Personal Services


```{r}
# Need to find the 5 NAICS codes with the most PPP total applications. 
#1), group the data by title
#2), find the sum of the applications for each industry
#3), sort the data in ascending order
loans_by_industry <- maryland_ppp_with_naics_and_demographics %>% 
  group_by(title) %>% 
  summarize(
    amount = n()) %>% 
arrange(desc(amount))
```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** 
1, Full-Service Restaurants --- $733008288
2, Offices of Physicians (except Mental Health Specialists) --- $407398101
3, Plumbing, Heating, and Air-Conditioning Contractors --- $324381106
4, Offices of Lawyers --- $291412051
5, Commercial and Institutional Building Construction --- $287903012

```{r}
# I need to find the SUM of the loan amounts per industry
# Then I need to filter to find the loan amounts totaling > $150,000
loan_sums_by_industry <- maryland_ppp_with_naics_and_demographics %>%
  group_by(title) %>% 
  summarise(
    total_loans_amount = sum(amount)) %>% 
  filter(total_loans_amount  > 15000) %>% 
  arrange(desc(total_loans_amount))

```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** 
Top zipcode: 20850
Count of loans: 3152	
Total loans amount: $473,471,466.53	
```{r}
# Second, I need to filter the data to be (PNHA >= 20)
# Third, I need to group by zipcode
# Fourth, I need to find the sum of the PPP loans for each zipcode and display the number of loans 
# Fifth, I need to arrange the data in descending order
maryland_ppp_with_naics_and_demographics %>%
  filter(PNHA >= 20) %>%
  group_by(zip5) %>%
  summarize(
    count = n(),
    total_loans_amount = sum(amount)) %>% 
    arrange(desc(total_loans_amount))


```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4**
Top zipcode: 20850
Count of loans: 239	
Total loans amount: $28,983,478.00	

```{r}
# Write the code to produce your answer here
#Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian",
maryland_ppp_with_naics_and_demographics %>%
  filter(race == "Asian") %>%
  group_by(zip5, PNHA) %>%
  summarize(
    count = n(),
    total_loans_amount = sum(amount)) %>% 
    arrange(desc(total_loans_amount))
```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants? 

**A5**
One possible explanation is that the applicants did not self-report (especially in the first draw), which would create a big gulf in the data to begin with. 


**Q6** Check to see if any Maryland zip codes that have loans where the PPP data lists the applicant as being Asian BUT are located in a zip code that does not appear to have Asian population according to the ZCTA data. How many loans do they have and what is the total amount? Take a look at these zip codes - what could explain this?
**A6** 
The zipcodes are 20701 (Annapolis Junction, MD — 4 loans, $75170.5), 21240 (Millersville, MD—3 loans, $358674.0	), and 21031 (Hunt Valley, MD —6 loans $6819563.0). The one thing I can think of is that the loan borrowers didn't self-report, or the lenders didn't keep track of the information.

```{r}
# Write the code to produce your answer here
maryland_ppp_with_naics_and_demographics %>%
  filter(race == "Asian") %>%
  filter(PNHA == 0) %>% 
  group_by(zip5, PNHA) %>%
  summarize(
    count = n(),
    total_loans_amount = sum(amount)) %>% 
    arrange((total_loans_amount))
```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** 
There are several, but I'd be interested to know why a business in the "Monetary Authorities-Central Bank" industry needed a loan. There are also others that seem unlikely loan-seekers, like businesses in the "Nuclear Electric Power Generation," "Small Arms Ammunition Manufacturing," "Guided Missile and Space Vehicle Manufacturing," "Industrial Gas Manufacturing," "Industrial Process Furnace and Oven Manufacturing," and "Ammunition (except Small Arms) Manufacturing." 
  
```{r}
# Write the code to produce your answer here
maryland_ppp_with_naics_and_demographics %>%
  group_by(title) %>% 
    summarize(
      count = n()) %>% 
  filter(count == 1)
  
```
